Stateless Model Checking with Data-Race Preemption Points

### 2016-03-24 oopsla-16/paper.tex
Write expanded explanation of HB for background section.
Expand related work to address PLDI reviewer concerns.
Fix miscounting in QS bugs, 163->157; and make ICB bug total the new denominator for claimed x-factor
Garth's comments thru soundness
Last of garth's commence

\usepackage{amsmath}
\usepackage{url}
\usepackage{enumitem}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{courier} % TODO: remove this? and get nicer fonts?
\usepackage{amsthm}
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}

@@

% TODO: remove all changebars before submission
%\setcounter{changebargrey}{0}
%\newcommand\revision[1]{\begin{samepage}\cbstart#1\cbend\end{samepage}}
%\newcommand\revision[1]{\cbstart#1\cbend}
%\newcommand\revision[1]{#1}
\definecolor{purple}{RGB}{192,0,192}
\newcommand\revision[1]{\hilight{purple}{#1}}

@@

\revision{We present \quicksand, a new stateless model checking framework}~which
manages the exploration of many state spaces using different preemption points.
It uses state space estimation to prioritize jobs most likely to complete in a fixed CPU budget,
and it incorporates data-race analysis to add new preemption points on the fly.
Preempting threads during a data race's instructions
%These preemption points allow us to check whether each data-race can produce an observable failure, and
can automatically classify the race as buggy or benign,
and uncovers new bugs not reachable by prior model checkers.
%
It also enables full verification of all possible schedules when every data race is verified as benign within the CPU budget.
In our evaluation, Iterative Deepening
%with data-race preemption points
found 1.8x as many bugs and verified 4.3x as many tests compared to stateless model checking without data-race preemption points.
%Our evaluation shows this technique is
%more effective than single-state-space model checking
%%both at finding more bugs and at completing more state spaces when no bug exists.
%both at finding the same bugs faster and at finding new bugs entirely.
\end{abstract}

@@

We achieve better bug-finding results than either single-pass data-race detection or single-state-space model checking alone,
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
%Compared to prior work,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
finding new bugs with data-race PPs that could not be exposed by preempting only on synchronization APIs.
By using \numstudence~student projects as our test suite, we also show the potential benefit as a debugging platform in educational settings.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide a verification as strong as if every single instruction had been used as a PP.

>>>>

\usepackage{amsmath}
\usepackage{url}
\usepackage{enumitem}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}
\newcommand{\doi}[1]{doi:~\href{http://dx.doi.org/#1}{\Hurl{#1}}}

@@

% TODO: remove all changebars before submission
%\setcounter{changebargrey}{0}
%\newcommand\revision[1]{\begin{samepage}\cbstart#1\cbend\end{samepage}}
%\newcommand\revision[1]{\cbstart#1\cbend}
\newcommand\revision[1]{#1}
\definecolor{purple}{RGB}{192,0,192}
%\newcommand\revision[1]{\hilight{purple}{#1}}

@@

\revision{We present \quicksand, a new stateless model checking framework}~which
manages the exploration of many state spaces using different preemption points.
It uses state space estimation to prioritize jobs most likely to complete in a fixed CPU budget,
and it incorporates data-race analysis to add new preemption points on the fly.
Preempting threads during a data race's instructions
%These preemption points allow us to check whether each data-race can produce an observable failure, and
can automatically classify the race as buggy or benign,
and uncovers new bugs not reachable by prior model checkers.
%
It also enables full verification of all possible schedules when every data race is verified as benign within the CPU budget.
In our evaluation, \quicksand~found
%with data-race preemption points
1.5x as many bugs and verified 4.3x as many tests compared to \revision{prior model checking approaches}. %stateless model checking without data-race preemption points.
%Our evaluation shows this technique is
%more effective than single-state-space model checking
%%both at finding more bugs and at completing more state spaces when no bug exists.
%both at finding the same bugs faster and at finding new bugs entirely.
\end{abstract}

@@

We achieve better bug-finding results than either single-pass data-race detection or single-state-space model checking alone,
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
%Compared to prior work,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
finding new bugs with data-race PPs that could not be exposed by preempting only on synchronization APIs.
By using \numstudence~student operating system implementation projects as our test suite, we also show the potential benefit as a debugging platform in educational settings.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide a verification as strong as if every single instruction had been used as a PP.

### 2016-03-24 oopsla-16/related.tex
Garth's comments thru soundness
Last of garth's commence

\revision{{\bf Restricting preemptions.}
Preemption Sealing \cite{sealing} pre\-sents a mechanism similar to the {\tt within\_function} command (\sect{\ref{sec:landslide}}) for users to manually restrict preemptions.
It demonstrates the need to consider subsets of PPs,
as well as developers' willingness to limit a test case's scope so the resulting state space may be fully verified (\sect{\ref{sec:totalverif}}).
%It justifies our assumption that expert
%programmers can customize small test cases which, when fully verified,
%provide a correctness assurance much stronger than the assurance from
%a partially-complete exploration of a huge state space.
Probabilistic Concurrency Testing (PCT) \cite{randomized-scheduler}
is a randomized algorithm that can quantify the probability of uncovering bugs.
PCT is optimized for tests with impossibly large state spaces,
while for more approachable tests,
its incompatibility with DPOR would significantly delay completion.
Future work could use ETAs to heuristically switch between DPOR and PCT}.
Finally, Iterative Context Bounding (ICB) \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB is limited to a fixed set of PPs,
\revision{cannot estimate remaining time to total verification,
and cannot incorporate data-race PPs not statically coded in advance}.
%and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

@@

%Too many related projects to list have made contributions to the
Many advances have been made on the false-positive data race problem since it was first introduced in \cite{eraser}.
\cite{hybriddatarace} and \cite{tsan}
%combine the lockset and happens-before analyses into a hybrid technique,
\revision{introduce Limited Happens-Before}, %(\sect{\ref{sec:overview-dr}})},
which we employ,
\revision{while \cite{predictive-dr} introduces the Causally-Precedes relation for a sound yet incomplete analysis (\sect{\ref{sec:overview-dr}})}.
\revision{Several tools \cite{droidracer,cafa,unnamed-android} have recently emerged to target Android applications},
%DroidRacer \cite{droidracer} and CAFA \cite{cafa} target
%%event-driven
%Android applications,
using domain-specific heuristics (orthogonal to our method) to reduce false positives. % cut for space?
Like \landslide, DataCollider \cite{datacollider} \revision{finds data races in}~kernel code.
IFRit \cite{ifrit}
improves performance using an interference analysis,
which could allow future work to avoid tracing every memory access.
%although in our context, we cannot admit false negatives (\sect{\ref{sec:totalverif}}).
% No, IDGAF about pure happens before.
%FastTrack \cite{fasttrack} optimizes the performance of pure happens-

>>>>

\revision{{\bf Restricting preemptions.}
Preemption Sealing \cite{sealing} pre\-sents a mechanism similar to the {\tt within\_function} command (\sect{\ref{sec:landslide}}) for users to manually restrict preemptions.
It demonstrates the need to consider subsets of PPs,
as well as developers' willingness to limit a test case's scope so the resulting state space may be fully verified (\sect{\ref{sec:totalverif}}).
%It justifies our assumption that expert
%programmers can customize small test cases which, when fully verified,
%provide a correctness assurance much stronger than the assurance from
%a partially-complete exploration of a huge state space.
Probabilistic Concurrency Testing (PCT) \cite{randomized-scheduler}
is a randomized algorithm that can quantify the probability of uncovering bugs.
PCT is optimized for tests with impossibly large state spaces,
while for more approachable tests,
its incompatibility with DPOR would significantly delay completion.
Future work could use ETAs to heuristically switch between DPOR and PCT}.
Finally, ICB \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB is limited to a fixed set of PPs,
\revision{cannot estimate remaining time to total verification,
and cannot incorporate data-race PPs not statically coded in advance}.
%and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

@@

%Too many related projects to list have made contributions to the
Many advances have been made on the false-positive potential data race problem since it was first introduced in \cite{eraser}.
\cite{hybriddatarace} and \cite{tsan}
%combine the lockset and happens-before analyses into a hybrid technique,
\revision{introduce Limited Happens-Before}, %(\sect{\ref{sec:overview-dr}})},
which we employ,
\revision{while \cite{predictive-dr} introduces the Causally-Precedes relation for a sound yet incomplete analysis (\sect{\ref{sec:overview-dr}})}.

\revision{Several tools \cite{droidracer,cafa,unnamed-android} have recently emerged to target Android applications},
%DroidRacer \cite{droidracer} and CAFA \cite{cafa} target
%%event-driven
%Android applications,
using domain-specific heuristics (orthogonal to our method) to reduce false positives. % cut for space?
Like \landslide, DataCollider \cite{datacollider} \revision{finds data races in}~kernel code.
IFRit \cite{ifrit}
improves the performance of HB using an interference analysis,
which could allow future work to avoid tracing every memory access.
%although in our context, we cannot admit false negatives (\sect{\ref{sec:totalverif}}).
% No, IDGAF about pure happens before.
%FastTrack \cite{fasttrack} optimizes the performance of pure happens-
